Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-07-11 prop_letter.tex
prop_letter.tex Updating...

There are long-standing debates in the foundations of quantum
mechanics regarding the tension between finite precision measurement
and contextuality\nb\cite{BarrettKent2004,Appleby_2005}.  For example,
Meyer\nb\cite{PhysRevLett.83.3751} and Mermin\nb\cite{Mermin1999} disagree
on whether finite-precision measurements invalidate the Kochen-Specker
theorem\nb\cite{kochenspecker1967,peres1995quantum}. In particular,
the Kochen-Specker theorem is based on the two sphere\nb$S^{2}$
cannot be colored by red and green such that any orthogonal basis
has exactly one red vector. On the one hand, Meyer argued that the
rational dense subset of $S^{2}$, $S^{2}\cap\mathbb{Q}^{3}$, is
colorable, and we cannot distinguish $S^{2}$ and $S^{2}\cap\mathbb{Q}^{3}$
since finite precision measurement\nb\citep{PhysRevLett.83.3751,Szangolies2015}.
On the another hand, Mermin argued any physical outcome must be a
continuous function on the experimental configuration\nb\citep{Mermin1999,Szangolies2015}.
Discussion
has continued on whether contextuality is consistent with  experimental outcomes\nb\cite{Spekkens2005,GuehneKleinmannCabelloEtAl2010,MazurekPuseyKunjwalEtAl2016}. The
fundamental question is how to make mathematical theories of quantum mechanics that are intrinsically, rather than only implicitly, consistent with the resources available for the realistic accuracy of an actual measurement.

@@

% exploiting the fact that a density matrix approach can be
% considered as equivalent to a description of quantum states
% using probability  measures\nb\cite{10.2307/2308516,Varadarajan2008}. %Mackey=10.2307

In particular, the effect of finite precision on contextuality and the
Kochen-Specker theorem is shown to have a resolution in terms of our quantum interval-valued probability measure framework.  To proceed, we use a formulation of the Mermin square that uses the
expectation value of an observable with respect
to a QIVPM can be defined to be consistent with the constraints on the
expectation value of the same observable expected from continuous
quantum mechanics.  We then examine Gleason's
theorem\nb\cite{gleason1957,Redhead1987-REDINA,peres1995quantum} and the
Kochen-Specker theorem in the context of QIVPM's, and conclude that
they correspond to the limit of an infinite number of precise
intervals, and the set of cryptodeterministic intervals (the set\nb$\{\imposs,\necess\}$),
respectively.

${\bullet}$ A key aspect of quantum measurement is its stochastic nature as reflected in Born's rule.  Gleason's theorem establishes the existence of a unique state (in Hilbert spaces of dimension $\ge 3$), the probabilities of whose outcomes are computed from Born's rule assuming infinite precision.  Then it is legitimate to question what happens when infinite precision cannot be achieved.

${\bullet}$ Exploiting the fact that a density matrix approach can be
considered as equivalent to a description of quantum states using
probability measures\nb\cite{10.2307/2308516,Varadarajan2008}. %Mackey=10.2307

We note, however, that in this approach there exist QIVPM's that do
not correspond to a density matrix computable by the Born rule, so
that Gleason's theorem cannot hold in general.  It thus remains
possible that a refinement of the basic definition of a quantum state
or the computation of probability from a state may be required when we
rigorously consider the implications of finite precision measurement.

A question we want to entertain in this paper is \ldots
Can one incorporate in a rigorous mathematical framework the idea of
finite precision measurement? Yes QIVPM. Now what is the fate of the 
foundational theorems of QM when this framework is applied?

@@

It is now not difficult to verify that for any totally mixed
state\nb$\rho$ on a $d$-dimensional Hilbert space,
$\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$ is a
QIVPM. Furthemore, we have:
\begin{itemize}
\item If $\left\Vert \rho-\rho'\right\Vert \le1/\left(4dn\right)$, $\rho'\in\coreBorn\left(\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}},\events\right)$.
\item If $\rho'\in\coreBorn\left(\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}},\events\right)$,
then $\left\Vert \rho-\rho'\right\Vert \le1/n$.
\end{itemize}
In other words, the states consistent with this QIVMP
$\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$ will be closer and
closer to $\rho$ as $n$ goes to infinity. In the limiting case when
the intervals are infinitely precise, i.e.,
$\set{\left[x,x\right]}{x\in\left[0,1\right]}$, any QIVPM can be
recast as an real-valued one and will have a unique state consistent
with it on every projector when $d\ge3$.

>>>>

There are long-standing debates in the foundations of quantum
mechanics regarding the tension between finite precision measurement
and contextuality\nb\cite{BarrettKent2004,Appleby_2005}.  For example,
Meyer\nb\cite{PhysRevLett.83.3751} and Mermin\nb\cite{Mermin1999} disagree
on whether finite-precision measurements invalidate the Kochen-Specker
theorem\nb\cite{kochenspecker1967,peres1995quantum}. In particular,
the Kochen-Specker theorem is based on the two sphere\nb$S^{2}$
cannot be colored by red and green such that any orthogonal basis
has exactly one red vector. On the one hand, Meyer argued that the
rational dense subset of $S^{2}$, $S^{2}\cap\mathbb{Q}^{3}$, is
colorable, and we cannot distinguish $S^{2}$ and $S^{2}\cap\mathbb{Q}^{3}$
since finite precision measurement\nb\citep{PhysRevLett.83.3751,Szangolies2015}.
On the another hand, Mermin argued any physical outcome must be a
continuous function on the experimental configuration so that small
experimental imperfection should not ``nullify'' an idealized theorem\nb\citep{Mermin1999,Szangolies2015}.
Discussion
has continued on whether contextuality is consistent with  experimental outcomes\nb\cite{Spekkens2005,GuehneKleinmannCabelloEtAl2010,MazurekPuseyKunjwalEtAl2016}. The
fundamental question is how to make mathematical theories of quantum mechanics that are intrinsically, rather than only implicitly, consistent with the resources available for the realistic accuracy of an actual measurement.

@@

In particular, the effect of finite precision on contextuality and the
Kochen-Specker theorem is shown to have a resolution in terms of our quantum interval-valued probability measure framework.  To proceed, we use a formulation of the Mermin square that uses the
expectation value of an observable with respect
to a QIVPM, where the interval-valued
expectation value is consistent with the expectation value of the
same observable on continuous quantum mechanics. Furthermore, the
Kochen-Specker theorem is robust in our framework in a sense that
it would remain valid when the experimental error is less than a particular
bound.

A key aspect of quantum measurement is its stochastic nature as reflected
in Born's rule\nb\citep{Born1983,peres1995quantum,544199}. Gleason's
theorem establishes the existence of a unique state (in Hilbert spaces
of dimension $\ge3$), the probabilities of whose outcomes are computed
from Born's rule assuming infinite precision\nb\citep{gleason1957,Redhead1987-REDINA,peres1995quantum}.
Then it is legitimate to question what happens when infinite precision
cannot be achieved. While a QIVPM\nb$\bar{\mu}$ might not be consistent
with a unique state\nb$\rho'$ on all projectors on a Hilbert space,
if $\bar{\mu}$ is induced by a mixed state\nb$\rho$, $\rho'$ would
close to $\rho$ when we shirk the length of intervals under some
good conditions, and recover original Gleason's theorem in the limiting
case. Since Gleason's theorem cannot apply on QIVPMs in general, it
thus remains possible that a refinement of the basic definition of
a quantum state or the computation of probability from a state may
be required when we rigorously consider the implications of finite
precision measurement.

@@

It is now not difficult to verify that for any totally mixed
state\nb$\rho$ on a three-dimensional Hilbert space,
$\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$ is a
QIVPM. Furthemore, we have:
\begin{itemize}
\item If $\left\Vert \rho-\rho'\right\Vert \le1/\left(12n\right)$,
$\rho'\in\coreBorn\left(\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}},\events\right)$.
\item If $\rho'\in\coreBorn\left(\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}},\events\right)$,
then $\left\Vert \rho-\rho'\right\Vert \le1/n$.
\end{itemize}
In other words, the states consistent with this QIVMP
$\widehat{\iota}_{n}\circ\mu_{\rho}^{\mathrm{B}}$ will be closer and
closer to $\rho$ as $n$ goes to infinity. In the limiting case when
the intervals are infinitely precise, i.e.,
$\set{\left[x,x\right]}{x\in\left[0,1\right]}$, any QIVPM can be
recast as an real-valued one and will have a unique state consistent
with it on every projector when $d\ge3$.
